### 2022-09-06 latex/Chapters/DRLSearch.tex
update

Since my implementation of \textbf{DQL} is no longer just a cost-to-go estimate as with \textbf{DRL}, but a joint cost-to-go and distribution over actions (to be concrete, that means a 5-dimensional vector for the \textbf{SP} and a 13-dimensional vector for the \textbf{RC}). This leaves us with the question of how can we use the output of such a network to search for a solution. One simple answer (which I have tried, see e.g. subsection \ref{sec:S33DRLDQL}) is that we could disregard the distribution and only take into account the cost-to-go by simply using $A^{*}$. The rationale for doing so is that we might hope that jointly learning the cost-to-go and the actions via a combination of MSE and cross-entropy-loss produces a better network and leads to better heuristics, since we use more information during the learning and iterations.
\\
\\
Another possibility is to use an algorithm which combines both the actions distribution and the value function to inform its search. The Monte Carlo Tree Search (\textbf{MCTS}) is such an algorithm. In short, it generates at each step (and possibly in a distributed manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until the goal is found). The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (following the most promising actions) as well as the value function, according to a trade-off which can be tuned via a hyper-parameter. In order to prevent always following the same paths (which might take us very far from the goal), \textbf{MCTS} algorithms usually keep track of actions they have followed already and underweight their probability so that subsequent paths explore other actions. A pseudo code description of it looks like this:
\\
\teal
\paragraph{}{\textbf{pseudo code -- \textbf{MCTS}}}
\begin{pseudocode}
#################################################################
def MCTS(initial_node, time_out, q_v_network):
    """ pseudo code for Monte Carlo Tree Search """
    initialize_time_out(time_out)
    tree = Tree(initial_node)
    while True:
        check_time_out()
        leaf = construct_path_to_new_leaf(tree, q_v_network)
        if leaf.is_goal():
            return BFS(tree) # path to leaf can usually be improved by BFS
#################################################################
def construct_path_to_new_leaf(tree, q_v_network):
    leaf = tree.initial_node
    while leaf.expanded:
        actions_probas, actions_values = q_v_network(leaf)
        """ choose next node based on joint actions probas & values
        as well as history of actions taken
        """
        leaf = leaf.choose_child(actions_probas,
                                 actions_values)
    # add children of leaf to tree
    leaf.expand()
    # penalize path taken to favour exploration
    leaf.increment_historical_actions_count()
    return leaf
#################################################################
\end{pseudocode}
\black

>>>>

Since my implementation of \textbf{DQL} is no longer just a cost-to-go estimate as with \textbf{DRL}, but a joint cost-to-go and distribution over actions (to be concrete, that means a 5-dimensional vector for the \textbf{SP} and a 13-dimensional vector for the \textbf{RC}). This leaves us with the question of how can we use the output of such a network to search for a solution. One simple answer (which I have tried, see e.g. subsection \ref{sec:S33DRLDQL}) is that we could disregard the distribution and only take into account the cost-to-go by simply using $A^{*}$. The rationale for doing so is that we might hope that jointly learning the cost-to-go and the actions via a combination of MSE and cross-entropy-loss produces a better network and leads to better heuristics, since we use more information during the learning and iterations.
\\
\\
Another possibility is to use an algorithm which combines both the actions distribution and the value function to inform its search. The Monte Carlo Tree Search (\textbf{MCTS}) is such an algorithm. In short, it generates at each step (and possibly in a distributed or multithreaded manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until the goal is found). The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (following the most promising actions) as well as the value function, according to a trade-off which can be tuned via a hyper-parameter. In order to prevent always following the same paths (which might take us very far from the goal), \textbf{MCTS} algorithms usually keep track of actions they have followed already and underweight their probability so that subsequent paths explore other actions. A pseudo code description of it looks like this:
\\
\teal
\paragraph{}{\textbf{pseudo code -- \textbf{MCTS}}}
\begin{pseudocode}
#################################################################
def MCTS(initial_node, time_out, q_v_network):
    """ pseudo code for Monte Carlo Tree Search """
    initialize_time_out(time_out)
    tree = Tree(initial_node)
    while True:
        check_time_out()
        leaf = construct_path_to_new_leaf(tree, q_v_network)
        if leaf.is_goal():
            return BFS(tree) # path to leaf can usually be improved by BFS
#################################################################
def construct_path_to_new_leaf(tree, q_v_network):
    leaf = tree.initial_node
    while leaf.expanded:
        actions_probas, actions_values = q_v_network(leaf)
        """ choose next node based on joint actions probas & values
        as well as history of actions taken
        """
        leaf = leaf.choose_child(actions_probas,
                                 actions_values)
    # add children of leaf to tree
    leaf.expand()
    # penalize path taken to favour exploration
    leaf.increment_historical_actions_count()
    return leaf
#################################################################
\end{pseudocode}
\black

### 2022-09-06 latex/Chapters/Implementation.tex
update

\begin{itemize}
\item \textbf{DFSSolver} This solver is literally a few lines of code, extremely close to the pseudo code detailed in \ref{sec:TheoryBFSDFS}
\item \textbf{BFSSolver} Ditto, see \ref{sec:TheoryBFSDFS}
\item \textbf{AStarSolver} Ditto, see \ref{sec:TheoryAStar}
\item \textbf{NaiveSlidingSolver} The NaiveSolver is conceptually simple, but was far from trivial to implement and debug. Its functioning is detailed in the appendix section \ref{sec:Examples}. Basically, it does what a beginner player would do (certainly what I do when I play the sliding puzzle), solving the top row first, then the left column, hence reducing the puzzle to a smaller dimensional one, and keep iterating until done.
\item \textbf{KociembaSolver} This solver is a wrapper around two libraries. The first one, hkociemba (see \cite{HKociemba}) is a 2x2x2 implementation written by Kociemba himself, the second one is a Python pip-installable which solves the 3x3x3 Rubik's. My  wrapper simply translates my cube representation in that accepted by these 2 solvers, and their result string into moves usable by my code. I also do some additional massaging for the 3x3x3 case, due to a shortcoming/bug in the 3x3x3 implementation, as I have detailed in appendix \ref{sec:Kociemba}.
\item \textbf{MonteCarloSearchTreeSolver} TBI I want to implement this in August when working on the Rubiks' ... following the Rubik's paper fromMcAleer \& Agostilenni et al \cite{https://doi.org/10.48550/arxiv.1805.07470}
\end{itemize}

\section{rubiks.scripts}

\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebasescripts}
%\decoRule
\caption[Codebase]{rubiks.scripts}
\label{fig:Codebasescripts}
\end{figure}

Finally it is worth noting that the code will save on disk a lot of data (e.g. the learners will save what they have learnt, e.g. a Pytorch network or a data base of optimal costs, the performance comparison will run solvers versus very many configurations of puzzles and save the results for later being able to display) etc... The base of the tree to save all this data can be chosen by setting up the "RUBIKSDATA" environment variable. If not, it will go somewhere in you HOME :)

>>>>

\begin{itemize}
\item \textbf{DFSSolver} This solver is literally a few lines of code, extremely close to the pseudo code detailed in \ref{sec:TheoryBFSDFS}
\item \textbf{BFSSolver} Ditto, see \ref{sec:TheoryBFSDFS}
\item \textbf{AStarSolver} Ditto, see \ref{sec:TheoryAStar}
\item \textbf{NaiveSlidingSolver} The NaiveSolver is conceptually simple, but was far from trivial to implement and debug. Its functioning is detailed in the appendix section \ref{sec:Examples}. Basically, it does what a beginner player would do, certainly what I do when I play the sliding puzzle: solving the top row first, then the left column, hence reducing the puzzle to a smaller dimensional one, and keep iterating until done.
\item \textbf{KociembaSolver} This solver is a wrapper around two libraries. The first one, hkociemba (see \cite{HKociemba}) is a 2x2x2 implementation written by Kociemba himself, the second one is a Python pip-installable library which solves the 3x3x3 Rubik's. My  wrapper simply translates my cube representation in that accepted by these 2 solvers, and their result string into moves usable by my code. I also do some additional massaging for the 3x3x3 case, due to a shortcoming/bug in the 3x3x3 Python library, as I have detailed in appendix \ref{sec:Kociemba}.
\item \textbf{MonteCarloSearchTreeSolver} I have implemented the algorithm described in the paper from McAleer \& Agostilenni et al \cite{https://doi.org/10.48550/arxiv.1805.07470}. My implementation is single-threaded however (in part due to my using Python for this project, which would have made using multiprocessing not efficient at all for their algorithm). A consequence of that is that I dropped the $\nu$ parameter from their implementation, since it only introduces a penalty to discourage parallel threads from generating the same paths from initial node to leaves (in single threaded mode, this parameter is inconsequential since the generation of a path decrements it, only to reincrement it once the leaf is generated).
\\
This solver basically generates (pseudo-random) paths from initial node to a new leaf at each iteration, until we obtain a goal leaf. I use the word pseudo-random since the paths are actually completely deterministic. They indeed result from a trade-off between the probability of actions (penalised over time by discounting actions which have been taken already) and the value function, but there is no random draws performed anywhere in there.
\\ As is described in McAleer \& Agostilenni's paper, I have implemented a final pruning via BFS of the tree resulting from the above \textit{Monte-Carlo} process consisting in generating random paths until a goal state is reached. Indeed, it is clear that the paths generated have very little chance of being optimal since there is no correction mechanism to avoid long and unecessary loops. I have however made this tree trimming by BFS optional via a \textit{trim\_tree} parameter as I wanted to study its effect on solutions' optimality and run time (see results section \ref{sec:MTCSHyperParamsEffect}).
\\
Similarly to their implementation, the trade off between optimal actions and value function is driven by a hyper-parameter \textit{c}, as I wanted to study its effect as well, and confirm my intuitions as to how it would affect solutions' quality and run time (again see results section \ref{sec:MTCSHyperParamsEffect}).

\section{data}

Finally it is worth noting that the code can/will save on disk a lot of data, including:
\begin{itemize}
\item Manhattan++ generates databases of linear constraint penalties (see \ref{HSS})
\item TrainingData: (see e.g. \ref{DLSS}) generates databases of puzzles and their associated costs for later training of a DeepLearner.
\item Learners: models generated by the DeepLearner, DeepReinforcementLearner, DeepQLearner as well as some convergence data useful to understand how the networks learnt. Similarly PerfectLearner saves its results for later reuse (in e.g. PerfectHeuristic) or display.
\item Solver: Sequences of scrambled puzzles which we use to make the performance tests (comparing different solvers) fair.
\item Solver.performance\_test: Results of the performance tests (all the metrics discussed earlier in \ref{sec:codesolvers}, for each solver and each difficulty level)
\item etc \dots
\end{itemize}
The data is saved according to a file system tree of the form:
$$root/data\_type/puzzle\_type/dimension/file\_name.pkl$$
where the root is driven by the "RUBIKSDATA" environment variable or, if not set, will go somewhere in your HOME.
